---
title: "conditionals_loops"
format: html
editor: visual
---

```{r}
#installing packages
library(stringr)
library(switchr)
```

```{r}
pm2_5 <- 160

if (pm2_5 < 100) {
  print("Low to moderate risk")
} else if (pm2_5 >= 100 & pm2_5 < 150) {
  print("Unhealthy for sensitive groups")
} else
  print("Health risk present")

```

```{r}
species <- "blue whale"

if (str_detect(species, "blue whale")) {
  print("you found a whale")
}
```
# Store the base price of a burrito as base_burrito with a value of 6.50. Store main_ingredient with a starting string of “veggie.” Write a statement that will return the price of a burrito based on what a user specifies as “main_ingredient” (either “veggie”, “chicken” or “steak”) given the following:

A veggie burrito is the cost of a base burrito
A chicken burrito costs 3.00 more than a base burrito
A steak burrito costs 3.25 more than a base burrito

```{r}
base_burrito <- 6.50

main_ingredient <- "chicken"

chicken_burrito <- base_burrito + 3.00
steak_burrito <- base_burrito + 3.50

switch(main_ingredient,
          "veggie" = print(base_burrito),
          "chicken" = print(chicken_burrito),
          "steak" = print(steak_burrito))

```

# Task 4
Create a new vector called fish that contains the values 8, 10, 12, 23 representing counts of different fish types in a fish tank (goldfish, tetras, guppies, and mollies, respectively). Write a for loop that iterates through fish, and returns what proportion of total fish in the tank are that species. Assume that these counts represent all fish in the tank.

```{r}
#num_fish <- c(8,10,12,23)
#total_fish <-(8 + 10 +12 +23)
#type_fish <- c("goldfish", "tetras", "guppies", "mollies")
fish <- c(goldfish = 8, tetras = 10, guppies = 12, mollies = 23)

for (i in seq_along(fish)) {
  fish_prop <- fish[i] / sum(fish)
  print(fish_prop)
}
```
# Task 5
There is an existing vector in R called month.name that contains all month names (just ry running month.name in the Console to check it out). Write a for loop that iterates over all months in month.name and prints “January is month 1,” “February is month 2”, etc.

Hint: you can index values in the month.name vector just like you would any other vector (e.g., try running month.name[5]).

```{r}
month.name
```
```{r}
for (i in seq_along(month.name)) {
    print(paste(month.name[i], "is month", i))
}
```






